### 2021-07-06 rad-icfp-2021/purerad.tex
typos

The target language of our source-code transformation is an extension to the source language.
It is a higher-order language, as our purely functional reverse-mode introduces a continuation.
The set of scalar operations should also be closed under partial differentiation. 
In more detail, for every unary scalar operation !op1!, 
we assumed given an operator $\partial_1$!op1! whose semantics should be the derivative of !op1!.
For every binary operator !op2!, we assume given operators $\partial_1$!op2!, $\partial_2$!op2! 
representing respectively the first and second partial derivative of !op2!.

Similarly, the target language contains more array primitives which are used to define the reverse derivatives of array operations. 
Scan left !scanl! is similar to fold left but also stores all the intermediate results in a array, which it returns.
In the same vein, scan right !scanr! perform a fold left by reading the array from right to left and also storing 
the intermediate results in a array from right to left.

Finally, we add two new shift operator !shift1L!, !shift1R!. 
They take an array of size $n$, and respectively forget the first and the last element of the array.
These somewhat ad-hoc operators naturally show up when differentiating fold-like operators.

The types and terms are presented in Figure~\ref{fig:target_grammar}.
Our lambda abstractions take $n$ arguments as we are not interested here with partial applications. 
In fact, the lambda abstractions introduced by reverse-mode will be removed during partial evaluation 
and the notation with lambda abstractions having $n$ bound variables makes reading slightly easier.
We note that we don't actually need the full power of higher-order because we only use lambda abstractions over variables of ground types
and let expressions binding such lambda abstractions. We only need the target language to be second-order.
The type system for the extended grammar of the target language is presented in Figure~\ref{fig:target_typesystem}.

@@

The idea is that $\rho$ represents the return type of the derivative part, which should be !A$_{1} \times \ldots \times$ A$_n$! 
if we want the whole gradient of a term !e! in context !$\Gamma = \;$x$_{1}$:A$_{1}$,$\ldots$,x$_n$:A$_n$!.
The subscript $\Gamma$ denotes the current context, 
which is locally augmented for instance when differentiating a !let! rule.
For non-unary operations, we differentiate the arguments from left to right and add their derivatives to the current stack, 
which is modeled by the continuation $Y$. 
Importantly for performance, each continuation variable $Y$ is only used once.

@@

Following the insight from Section~\ref{subsec:insights}, 
we don't want to keep all the lambda abstractions as this is costly. 
Fortunately, the transformation is designed in such a way that we only have applications 
and we can use partial evaluation to do beta-reduction.

>>>>

The target language of our source-code transformation is an extension to the source language.
It is a higher-order language, as our purely functional reverse-mode introduces a continuation.
The set of scalar operations should also be closed under partial differentiation. 
In more detail, for every unary scalar operation !op1!, 
we assumed given an operator $\partial_1$!op1! whose semantics should be the derivative of !op1!.
For every binary operator !op2!, we assume given operators $\partial_1$!op2!, $\partial_2$!op2! 
respectively representing the first and second partial derivative of !op2!.

Similarly, the target language contains more array primitives which are used to define the reverse derivatives of array operations. 
Scan left !scanl! is similar to fold left but also stores all the intermediate results in an array, which it returns.
In the same vein, scan right !scanr! performs a fold left by reading the array from right to left and stores 
the intermediate results in an array from right to left.

Finally, we add two new shift operators !shift1L! and !shift1R!. 
They take an array of size $n$, and respectively forget the first and the last element of the array.
These somewhat ad-hoc operators naturally show up when differentiating fold-like operators.

The types and terms are presented in Figure~\ref{fig:target_grammar}.
Our lambda abstractions take $n$ arguments as we are not interested here with partial applications. 
In fact, the lambda abstractions introduced by reverse-mode will be removed during partial evaluation, 
and the notation with lambda abstractions having $n$ bound variables makes reading slightly easier.
We note that we don't actually need the full power of higher-order because we only use lambda abstractions over variables of ground types
and let expressions binding such lambda abstractions. We only need the target language to be second-order.
The type system for the extended grammar of the target language is presented in Figure~\ref{fig:target_typesystem}.

@@

The idea is that $\rho$ represents the return type of the derivative part, which should be !A$_{1} \times \ldots \times$ A$_n$! 
if we want the whole gradient of a term !e! in context !$\Gamma = \;$x$_{1}$:A$_{1}$,$\ldots$,x$_n$:A$_n$!.
The subscript $\Gamma$ denotes the current context, 
which is locally augmented, for instance when differentiating a !let! rule.
For non-unary operations, we differentiate the arguments from the left to the right and add their derivatives to the current stack, 
which is modeled by the continuation $Y$. 
Importantly for performance, each continuation variable $Y$ is only used once.

@@

Following the insight from Section~\ref{subsec:insights}, 
we don't want to keep all the lambda abstractions as this is costly. 
Fortunately, the transformation is designed in such a way that we only have applications, 
and we can use partial evaluation to do beta-reduction.
